---
title: "Computational Modeling - Assignment 1"
author: "Riccardo Fusaroli"
date: "07/01/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## In this assignment we learn how to assess rates from a binomial distribution, using the case of assessing your teachers' knowledge of CogSci

N.B. this markdown has 2 parts as it spans 2 weeks of teaching

### First part

You want to assess your teachers' knowledge of cognitive science. "These guys are a bunch of drama(turgist) queens, mindless philosophers, chattering communication people and Russian spies. Do they really know CogSci?", you think.

To keep things simple (your teachers should not be faced with too complicated things):
- You created a pool of equally challenging questions on CogSci
- Each question can be answered correctly or not (we don't allow partially correct answers, to make our life simpler).
- Knowledge of CogSci can be measured on a scale from 0 (negative knowledge, all answers wrong) through 0.5 (random chance) to 1 (awesome CogSci superpowers)

This is the data:
- Riccardo: 3 correct answers out of 6 questions
- Kristian: 2 correct answers out of 2 questions (then he gets bored)
- Daina: 160 correct answers out of 198 questions (Daina never gets bored)
- Mikkel: 66 correct answers out of 132 questions

```{r}
library(rethinking)
library(tidyverse)

```


Questions:

1. What's Riccardo's estimated knowledge of CogSci? What is the probability he knows more than chance (0.5) [try figuring this out. if you can't peek into chapters 3.1 and 3.2 and/or the slides]?

- First implement a grid approximation (hint check paragraph 2.4.1!) with a uniform prior, calculate the posterior and plot the results


```{r}
#Riccardo: 3 correct answers out of 6 questions

#define the grid
dens <- 50
p_grid <- seq(from = 0 , to = 1 , length.out = dens)

#first trying with a uniform prior
uniform.prior <- rep(1, dens ) #flat

#test the prior
dens(rbinom(1e4, 6, runif(1e4, 0, 1))) #looks fine 

#calculate the likelihood for riccardo
likelihood.r <- dbinom(3 , size = 6 , prob = p_grid )

#compute posterior 
unstd.posterior.r <- likelihood.r * uniform.prior

#standardize the posterior
posterior.r <- unstd.posterior.r / sum(unstd.posterior.r)

#draw the plot
d <- data.frame(grid = p_grid, posterior = posterior.r, prior = uniform.prior, likelihood = likelihood.r)

ggplot(d, aes(grid,posterior)) +  geom_point() +geom_line()+theme_classic()+  geom_line(aes(grid, prior/dens),color= 'red')+  xlab("Riccardo's Knowledge of CogSci")+ ylab("posterior probability")+ggtitle("Grid approximation using a uniform prior")



```


```



- Then implement a quadratic approximation (hint check paragraph 2.4.2!).
- N.B. for the rest of the exercise just keep using the grid approximation (we'll move to quadratic approximations in two classes)

```

```{r}
#quadratic approximation for Riccardo

#quap needs a provided formula (list of data), defining the probability of the data and the prior

ric.qa <- quap(
    alist(
        C ~ dbinom( C+W ,p) , #binomial likelihood #C and W stands for correb and wrong
        p ~ dunif(0,1) #uniform prior
    ) ,
    data=list(C=3,W=3))

# display summary of quadratic approximation
precis(ric.qa)

#assuming the posterior is Gaussian, it is maximized at 0.5, and its standard deviation is 0.2"

```


2. Estimate all the teachers' knowledge of CogSci. Who's best? Use grid approximation. Comment on the posteriors of Riccardo and Mikkel.
2a. Produce plots of the prior, and posterior for each teacher.


```{r}
#Kristian: 2 correct answers out of 2 questions (then he gets bored)

#test the prior
dens(rbinom(1e4, 2, runif(1e4, 0, 1))) #looks fine 

#calculate the likelihood for Kristian
likelihood.k <- dbinom(2 , size = 2 , prob = p_grid )

#compute posterior 
unstd.posterior.k <- likelihood.k * uniform.prior

#standardize the posterior
posterior.k <- unstd.posterior.k / sum(unstd.posterior.k)

#draw the plot
d.k <- data.frame(grid = p_grid, posterior = posterior.k, prior = uniform.prior, likelihood = likelihood.k)

ggplot(d.k, aes(grid,posterior.k)) +  geom_point() +geom_line()+theme_classic()+  geom_line(aes(grid, prior/dens),color= 'red')+  xlab("Kristian's knowledge of CogSci")+ ylab("posterior probability")+ggtitle("Grid approximation using a uniform prior")

```

```{r}
#Daina: 160 correct answers out of 198 questions 

#test the prior
dens(rbinom(1e4, 198, runif(1e4, 0, 1))) #looks fine 

#calculate the likelihood for Daina
likelihood.d <- dbinom(160 , size = 198 , prob = p_grid )

#compute posterior 
unstd.posterior.d <- likelihood.d * uniform.prior

#standardize the posterior
posterior.d <- unstd.posterior.d / sum(unstd.posterior.d)

#draw the plot
d.d <- data.frame(grid = p_grid, posterior = posterior.d, prior = uniform.prior, likelihood = likelihood.d)

ggplot(d.d, aes(grid,posterior.d)) +  geom_point() +geom_line()+theme_classic()+  geom_line(aes(grid, prior/dens),color= 'red')+  xlab("Daina's knowledge of CogSci")+ ylab("posterior probability")+ggtitle("Grid approximation using a uniform prior")
```


```{r}

#Mikkel: 66 correct answers out of 132 questions

#test the prior
dens(rbinom(1e4, 132, runif(1e4, 0, 1))) #looks fine 

#calculate the likelihood for riccardo
likelihood.m <- dbinom(66 , size = 132 , prob = p_grid )

#compute posterior 
unstd.posterior.m <- likelihood.m * uniform.prior

#standardize the posterior
posterior.m <- unstd.posterior.m / sum(unstd.posterior.m)

#draw the plot
d.m <- data.frame(grid = p_grid, posterior = posterior.m, prior = uniform.prior, likelihood = likelihood.m)

ggplot(d.m, aes(grid,posterior.m)) +  geom_point() +geom_line()+theme_classic()+  geom_line(aes(grid, prior/dens),color= 'red')+  xlab("Mikkel's knowledge of CogSci")+ ylab("posterior probability")+ggtitle("Grid approximation using a uniform prior")

```


3. Change the prior. Given your teachers have all CogSci jobs, you should start with a higher appreciation of their knowledge: the prior is a normal distribution with a mean of 0.8 and a standard deviation of 0.2. Do the results change (and if so how)?
3a. Produce plots of the prior and posterior for each teacher.

```{r}
#define new prior
new.prior <- dnorm(p_grid, 0.8, 0.2) #sensibly centered at chance

```

Riccardo: 3 correct answers out of 6 questions


```{r}


#testing the prior
dens(rbinom(1e4, 6, rnorm(1e4, 0.8, 0.2))) 

#Compute the likelihood at each value in grid
likelihood.r <- dbinom(3 , size = 6 , prob = p_grid )

#Compute the posterior (likelihood by prior)
unstd.posterior.r <- likelihood.r * new.prior

#Standardize the posterior (so it sums to 1)
posterior.r <- unstd.posterior.r / sum(unstd.posterior.r)

#Draw the plot
d.r <- data.frame(grid = p_grid, posterior = posterior.r, prior = new.prior, likelihood.r = likelihood.r)
ggplot(d.r, aes(grid,posterior.r)) +  geom_point() +geom_line()+theme_classic()+  geom_line(aes(grid, new.prior/dens),color= 'green')+xlab("Riccardo's knowledge of CogSci")+ ylab("posterior probability")+ggtitle("Grid approximation using the updated prior")
```

Kristian: 2 correct answers out of 2 questions (then he gets bored)

```{r}

#testing the prior
dens(rbinom(1e4, 6, rnorm(1e4, 0.8, 0.2))) 

#calculate the likelihood for riccardo
likelihood.k <- dbinom(2 , size = 2 , prob = p_grid )

#compute posterior 
unstd.posterior.k <- likelihood.k * new.prior

#standardize the posterior
posterior.k <- unstd.posterior.k / sum(unstd.posterior.k)

#draw the plot
d.k <- data.frame(grid = p_grid, posterior = posterior.k, prior = new.prior, likelihood = likelihood.k)

ggplot(d.k, aes(grid,posterior.k)) +  geom_point() +geom_line()+theme_classic()+  geom_line(aes(grid, prior/dens),color= 'red')+  xlab("Kristian's knowledge of CogSci")+ ylab("posterior probability")+ggtitle("Grid approximation using the updated prior")
```

- Daina: 160 correct answers out of 198 questions (Daina never gets bored)

```{r}
#testing the prior
dens(rbinom(1e4, 198, rnorm(1e4, 0.8, 0.2))) 

#Compute the likelihood at each value in grid
likelihood.d <- dbinom(160 , size = 198 , prob = p_grid )

#Compute the posterior (likelihood by prior)
unstd.posterior.d <- likelihood.d * new.prior

#Standardize the posterior (so it sums to 1)
posterior.d <- unstd.posterior.d / sum(unstd.posterior.d)

#Draw the plot
d.d <- data.frame(grid = p_grid, posterior = posterior.d, prior = new.prior, likelihood.d = likelihood.d)
ggplot(d.d, aes(grid,posterior.d)) +  geom_point() +geom_line()+theme_classic()+  geom_line(aes(grid, new.prior/dens),color= 'green')+xlab("Daina's knowledge of CogSci")+ ylab("posterior probability")+ggtitle("Grid approximation using the updated prior")
```

- Mikkel: 66 correct answers out of 132 questions

```{r}
#testing the prior
dens(rbinom(1e4, 132, rnorm(1e4, 0.8, 0.2))) 

#Compute the likelihood at each value in grid
likelihood.m <- dbinom(66 , size = 132 , prob = p_grid )

#Compute the posterior (likelihood by prior)
unstd.posterior.m <- likelihood.m * new.prior

#Standardize the posterior (so it sums to 1)
posterior.m <- unstd.posterior.m / sum(unstd.posterior.m)

#Draw the plot
d.m <- data.frame(grid = p_grid, posterior = posterior.m, prior = new.prior, likelihood.d = likelihood.m)
ggplot(d.m, aes(grid,posterior.m)) +  geom_point() +geom_line()+theme_classic()+  geom_line(aes(grid, new.prior/dens),color= 'green')+xlab("Mikkel's knowledge of CogSci")+ ylab("posterior probability")+ggtitle("Grid approximation using the updated prior")
```


4. You go back to your teachers and collect more data (multiply the previous numbers by 100). Calculate their knowledge with both a uniform prior and a normal prior with a mean of 0.8 and a standard deviation of 0.2. Do you still see a difference between the results? Why?

Riccardo 100 uniform
```{r}
#Compute the likelihood at each value in grid
likelihood.r <- dbinom(300 , size = 600 , prob = p_grid )

#Compute the posterior (likelihood by prior)
unstd.posterior.r <- likelihood.r * uniform.prior

#Standardize the posterior (so it sums to 1)
posterior.r <- unstd.posterior.r / sum(unstd.posterior.r)

#Draw the plot
d.r <- data.frame(grid = p_grid, posterior = posterior.r, prior = uniform.prior, likelihood.r = likelihood.r)
ggplot(d.r, aes(grid,posterior.r)) +  geom_point() +geom_line()+theme_classic()+  geom_line(aes(grid, uniform.prior/dens),color= 'green')+xlab("Riccardo's knowledge of CogSci")+ ylab("posterior probability")+ggtitle("Grid approximation using the uniform prior")
```

Riccardo 100 other prior
```{r}
#Compute the likelihood at each value in grid
likelihood.r <- dbinom(300 , size = 600 , prob = p_grid )

#Compute the posterior (likelihood by prior)
unstd.posterior.r <- likelihood.r * new.prior

#Standardize the posterior (so it sums to 1)
posterior.r <- unstd.posterior.r / sum(unstd.posterior.r)

#Draw the plot
d.r <- data.frame(grid = p_grid, posterior = posterior.r, prior = new.prior, likelihood.r = likelihood.r)
ggplot(d.r, aes(grid,posterior.r)) +  geom_point() +geom_line()+theme_classic()+  geom_line(aes(grid, new.prior/dens),color= 'green')+xlab("Riccardo's knowledge of CogSci")+ ylab("posterior probability")+ggtitle("Grid approximation using the updated prior")

```

Kristian 100 uniform

```{r}
#Compute the likelihood at each value in grid
likelihood.k <- dbinom(200 , size = 200 , prob = p_grid )

#compute posterior 
unstd.posterior.k <- likelihood.k * uniform.prior

#standardize the posterior
posterior.k <- unstd.posterior.k / sum(unstd.posterior.m)

#draw the plot
d.k <- data.frame(grid = p_grid, posterior = posterior.k, prior = uniform.prior, likelihood = likelihood.k)

ggplot(d.k, aes(grid,posterior.k)) +  geom_point() +geom_line()+theme_classic()+  geom_line(aes(grid, prior/dens),color= 'red')+  xlab("Kristian's knowledge of CogSci")+ ylab("posterior probability")+ggtitle("Grid approximation using the uniform prior")
```
Kristian 100 other prior

```{r}
#Compute the likelihood at each value in grid
likelihood.k <- dbinom(200 , size = 200 , prob = p_grid )

#compute posterior 
unstd.posterior.k <- likelihood.k * new.prior

#standardize the posterior
posterior.k <- unstd.posterior.k / sum(unstd.posterior.m)

#draw the plot
d.k <- data.frame(grid = p_grid, posterior = posterior.k, prior = new.prior, likelihood = likelihood.k)

ggplot(d.k, aes(grid,posterior.k)) +  geom_point() +geom_line()+theme_classic()+  geom_line(aes(grid, prior/dens),color= 'red')+  xlab("Kristian's knowledge of CogSci")+ ylab("posterior probability")+ggtitle("Grid approximation using the updated prior")
```

Daina 100 uniform

```{r}
#Compute the likelihood at each value in grid
likelihood.d <- dbinom(16000 , size = 19800 , prob = p_grid )

#Compute the posterior (likelihood by prior)
unstd.posterior.d <- likelihood.d * uniform.prior

#Standardize the posterior (so it sums to 1)
posterior.d <- unstd.posterior.d / sum(unstd.posterior.d)

#Draw the plot
d.d <- data.frame(grid = p_grid, posterior = posterior.d, prior = uniform.prior, likelihood.d = likelihood.d)
ggplot(d.d, aes(grid,posterior.d)) +  geom_point() +geom_line()+theme_classic()+  geom_line(aes(grid, prior/dens),color= 'green')+xlab("Daina's knowledge of CogSci")+ ylab("posterior probability")+ggtitle("Grid approximation using the uniform prior")
```

Daina 100 other prior
```{r}
#Compute the likelihood at each value in grid
likelihood.d <- dbinom(16000 , size = 19800 , prob = p_grid )

#Compute the posterior (likelihood by prior)
unstd.posterior.d <- likelihood.d * new.prior

#Standardize the posterior (so it sums to 1)
posterior.d <- unstd.posterior.d / sum(unstd.posterior.d)

#Draw the plot
d.d <- data.frame(grid = p_grid, posterior = posterior.d, prior = new.prior, likelihood.d = likelihood.d)
ggplot(d.d, aes(grid,posterior.d)) +  geom_point() +geom_line()+theme_classic()+  geom_line(aes(grid, new.prior/dens),color= 'green')+xlab("Daina's knowledge of CogSci")+ ylab("posterior probability")+ggtitle("Grid approximation using the updated prior")
```

Mikkel 100 uniform


```{r}
#Compute the likelihood at each value in grid
likelihood.m <- dbinom(6600 , size = 13200 , prob = p_grid )

#Compute the posterior (likelihood by prior)
unstd.posterior.m <- likelihood.m * uniform.prior

#Standardize the posterior (so it sums to 1)
posterior.m <- unstd.posterior.m / sum(unstd.posterior.m)

#Draw the plot
d.m <- data.frame(grid = p_grid, posterior = posterior.m, prior = uniform.prior, likelihood.m = likelihood.m)
ggplot(d.m, aes(grid,posterior.m)) +  geom_point() +geom_line()+theme_classic()+  geom_line(aes(grid, prior/dens),color= 'green')+xlab("Mikkel's knowledge of CogSci")+ ylab("posterior probability")+ggtitle("Grid approximation using the uniform prior")
```
Mikkel 100 other prior

```{r}
#Compute the likelihood at each value in grid
likelihood.m <- dbinom(6600 , size = 13200 , prob = p_grid )

#Compute the posterior (likelihood by prior)
unstd.posterior.m <- likelihood.m * new.prior

#Standardize the posterior (so it sums to 1)
posterior.m <- unstd.posterior.m / sum(unstd.posterior.m)

#Draw the plot
d.m <- data.frame(grid = p_grid, posterior = posterior.m, prior = new.prior, likelihood.m = likelihood.m)
ggplot(d.m, aes(grid,posterior.m)) +  geom_point() +geom_line()+theme_classic()+  geom_line(aes(grid, new.prior/dens),color= 'green')+xlab("Mikkel's knowledge of CogSci")+ ylab("posterior probability")+ggtitle("Grid approximation using the updated prior")
```


5. Imagine you're a skeptic and think your teachers do not know anything about CogSci, given the content of their classes. How would you operationalize that belief?

6. Optional question: Can you estimate the difference between Riccardo's estimated knowledge and that of each of the other teachers? Would you deem it credible (that is, would you believe that it is actually different)? 

### Second part: Focusing on predictions

Last year you assessed the teachers (darned time runs quick!). Now you want to re-test them and assess whether your models are producing reliable predictions. In Methods 3 we learned how to do machine-learning style assessment of predictions (e.g. rmse on testing datasets). Bayesian stats makes things a bit more complicated. So we'll try out how that works. N.B. You can choose which prior to use for the analysis of last year's data.

Questions to be answered (but see guidance below):
1- Write a paragraph discussing how assessment of prediction performance is different in Bayesian vs. frequentist models

2- Provide at least one plot and one written line discussing prediction errors for each of the teachers.

This is the old data:
- Riccardo: 3 correct answers out of 6 questions
- Kristian: 2 correct answers out of 2 questions (then he gets bored)
- Daina: 160 correct answers out of 198 questions (Daina never gets bored)
- Mikkel: 66 correct answers out of 132 questions

This is the new data:
- Riccardo: 9 correct answers out of 10 questions (then he freaks out about teaching preparation and leaves)
- Kristian: 8 correct answers out of 12 questions
- Daina: 148 correct answers out of 172 questions (again, Daina never gets bored)
- Mikkel: 34 correct answers out of 65 questions

```{r}
#sampling from the posteriors for each teacher
set.seed(100)

#riccardo
samples.riccardo <- 
    d.r %>% 
    slice_sample(n = 1000, weight_by = d.r$posterior, replace = T)

#kristian
samples.kristian <- 
    d.k %>% 
    slice_sample(n = 1000, weight_by = d.k$posterior, replace = T)

#daina
samples.daina <- 
    d.d %>% 
    slice_sample(n = 1000, weight_by = d.d$posterior, replace = T)

#mikkel
samples.mikkel <- 
    d.m %>% 
    slice_sample(n = 1000, weight_by = d.m$posterior, replace = T)

```


```{r}
#doing predictions for Riccardo
prediction.r <- rbinom(n = 1e4, size = 10, prob = rnorm(1e4, mean(samples.riccardo$grid), sd(samples.riccardo$grid) ))

#simple plot
hist(prediction.r) 

#doing predictions for Kristian
prediction.k <- rbinom(n = 1e4, size = 12, prob = rnorm(1e4, mean(samples.kristian$grid), sd(samples.kristian$grid) ))

#simple plot
hist(prediction.k)

#doing predictions for Daina
prediction.d <- rbinom(n = 1e4, size = 172, prob = rnorm(1e4, mean(samples.daina$grid), sd(samples.daina$grid) ))

#simple plot
hist(prediction.d)

#doing predictions for Mikkel
prediction.m <- rbinom(n = 1e4, size = 65, prob = rnorm(1e4, mean(samples.mikkel$grid), sd(samples.mikkel$grid) ))

#simple plot
hist(prediction.m)


```

Now plotting the predictions and highlighting expectations using different colors

```{r}
#saving predictions as a dataframe 
prediction.r <- as.data.frame(prediction.r) #riccardo
prediction.k <- as.data.frame(prediction.k) #kristian
prediction.d <- as.data.frame(prediction.d) #daina
prediction.m <- as.data.frame(prediction.m) #mikkel

#renaming the coloumn so it is more meaningful
prediction.r <-  prediction.r %>% 
  rename(no.cor.answer = prediction.r) #riccardo

prediction.k <-  prediction.k %>% 
  rename(no.cor.answer = prediction.k) #kristian

prediction.d <-  prediction.d %>% 
  rename(no.cor.answer = prediction.d) #daina

prediction.m <-  prediction.m %>% 
  rename(no.cor.answer = prediction.m) #mikkel

#plotting Riccardo
ggplot(prediction.r, aes(x = no.cor.answer))+ #no. of correct answer on x-axis
    geom_histogram(aes(fill = no.cor.answer == 9), #histogram, separately coloring the actual no. of correct answers from new data to be able to compare with the predicted values 
                 binwidth = 1, center = 0,
                 color = "black" , size = 1/10) +  
    scale_fill_manual(values = c("gray60", "firebrick")) + #defining colors
    scale_x_continuous("Number of correct answers", #making the x-axis nice
                     breaks = seq(from = 0, to = 10, by = 1)) +
    ylab("Predicted count for Riccardo") +
    coord_cartesian(xlim = c(0, 10)) +
    theme_minimal() +
    theme(legend.position = "none")

#plotting Kristian
ggplot(prediction.k, aes(x = no.cor.answer))+
    geom_histogram(aes(fill = no.cor.answer == 8),
                 binwidth = 1, center = 0,
                 color = "black" , size = 1/10) +  
    scale_fill_manual(values = c("gray60", "firebrick")) +
    scale_x_continuous("Number of correct answers",
                     breaks = seq(from = 0, to = 12, by = 1)) +
    ylab("Predicted count for Kristian") +
    coord_cartesian(xlim = c(0, 12)) +
    theme_minimal() +
    theme(legend.position = "none")

#plotting Daina
ggplot(prediction.d, aes(x = no.cor.answer))+
    geom_histogram(aes(fill = no.cor.answer == 148),
                 binwidth = 1, center = 0,
                 color = "black" , size = 1/10) +  
    scale_fill_manual(values = c("gray60", "firebrick")) +
    scale_x_continuous("Number of correct answers",
                     breaks = seq(from = 0, to = 172, by = 10)) +
    ylab("Predicted count for Daina") +
    theme_minimal() +
    theme(legend.position = "none")

#plotting Mikkel
ggplot(prediction.m, aes(x = no.cor.answer))+
    geom_histogram(aes(fill = no.cor.answer == 34),
                 binwidth = 1, center = 0,
                 color = "black", size = 1/10) +  
    scale_fill_manual(values = c("gray60", "firebrick")) +
    scale_x_continuous("Number of correct answers",
                     breaks = seq(from = 0, to = 65, by = 5)) +
    ylab("Predicted count for Mikkel") +
    theme_minimal() +
    theme(legend.position = "none")
```

Other method, using previous posterior as new updated prior

```{r}
#Riccardo
new.r <- 
    d.r %>% 
    mutate(prior = posterior)

#Compute the likelihood at each value in grid
likelihood.r <- dbinom(9 , size = 10 , prob = p_grid )

#Compute the posterior (likelihood by prior)
unstd.posterior.r <- likelihood.r * new.r$prior

#Standardize the posterior (so it sums to 1)
posterior.r <- unstd.posterior.r / sum(unstd.posterior.r)

#Draw the plot
ggplot(new.r, aes(grid,posterior.r)) +  geom_point() +geom_line()+theme_classic()+  geom_line(aes(grid, prior),color= 'green')+xlab("Riccardo's knowledge of CogSci")+ ylab("posterior probability")+ggtitle("Grid approximation using the previous posterior as updated prior")

#kristian
new.k <- 
    d.k %>% 
    mutate(prior = posterior)

#Compute the likelihood at each value in grid
likelihood.k <- dbinom(8 , size = 12 , prob = p_grid )

#Compute the posterior (likelihood by prior)
unstd.posterior.k <- likelihood.k * new.k$prior

#Standardize the posterior (so it sums to 1)
posterior.k <- unstd.posterior.k / sum(unstd.posterior.k)

#Draw the plot
ggplot(new.k, aes(grid,posterior.k)) +  geom_point() +geom_line()+theme_classic()+  geom_line(aes(grid, prior),color= 'green')+xlab("Kristian's knowledge of CogSci")+ ylab("posterior probability")+ggtitle("Grid approximation using the previous posterior as updated prior")

#Daina
new.d <- 
    d.d %>% 
    mutate(prior = posterior)

#Compute the likelihood at each value in grid
likelihood.d <- dbinom(148 , size = 172 , prob = p_grid )

#Compute the posterior (likelihood by prior)
unstd.posterior.d <- likelihood.d * new.d$prior

#Standardize the posterior (so it sums to 1)
posterior.d <- unstd.posterior.d / sum(unstd.posterior.d)

#Draw the plot
ggplot(new.d, aes(grid,posterior.d)) +  geom_point() +geom_line()+theme_classic()+  geom_line(aes(grid, prior),color= 'green')+xlab("Daina's knowledge of CogSci")+ ylab("posterior probability")+ggtitle("Grid approximation using the previous posterior as updated prior")

#Mikkel
new.m <- 
    d.m %>% 
    mutate(prior = posterior)

#Compute the likelihood at each value in grid
likelihood.m <- dbinom(34 , size = 65 , prob = p_grid )

#Compute the posterior (likelihood by prior)
unstd.posterior.m <- likelihood.m * new.m$prior

#Standardize the posterior (so it sums to 1)
posterior.m <- unstd.posterior.m / sum(unstd.posterior.m)

#Draw the plot
ggplot(new.m, aes(grid,posterior.m)) +  geom_point() +geom_line()+theme_classic()+  geom_line(aes(grid, prior),color= 'green')+xlab("Mikkel's knowledge of CogSci")+ ylab("posterior probability")+ggtitle("Grid approximation using the previous posterior as updated prior")
```

When comparing the updated prior (green line) to the posterior distribution for each teacher, we can highlight the errors of the prediction. 

Highest posterior density intervals:
```{r}


#check HPDI for Riccardo
samples.r<-sample(p_grid, prob=posterior.r, size=1e4, replace=TRUE)
HPDI(samples.r, 0.5)

#check HPDI for Kristian
samples.k<-sample(p_grid, prob=posterior.k, size=1e4, replace=TRUE)
HPDI(samples.k, 0.5)

#check HPDI for Daina
samples.d<-sample(p_grid, prob=posterior.d, size=1e4, replace=TRUE)
HPDI(samples.d, 0.5)

#check HPDI for Mikkel
samples.m<-sample(p_grid, prob=posterior.m, size=1e4, replace=TRUE)
HPDI(samples.m, 0.5)


```

And plotting the above intervals:

```{r}

#plot the HPDI results
p2 <-
  new.r %>% 
  ggplot(aes(x = p_grid)) +
  geom_line(aes(y = posterior.r)) +
  geom_ribbon(data = new.r %>% filter(p_grid > 0.67 & p_grid < 0.71),
              aes(ymin = 0, ymax = posterior.r)) +
  annotate(geom = "text",
           x = .25, y = .0025,
           label = "HPDI") +
  labs(x = "Proportion of correct questions (p)",
       y = "density")

p2




p <-
  new.r %>% 
  ggplot(aes(x = p_grid)) +
  geom_ribbon(data = . %>% filter(p_grid > 0.69 & p_grid < 0.71),
              aes(ymin = 0, ymax = posterior),
              fill = "grey75") +
  geom_line(aes(y = posterior)) +
  labs(subtitle = "50% HPDI",
       x = "proportion of water (p)",
       y = "density")

p
```


Guidance Tips

1. There are at least two ways of assessing predictions.
2. Last year's results are this year's expectations.
3. Are the parameter estimates changing? (way 1)
4. How does the new data look in last year's predictive posterior? (way 2)

When implementing the other way of assessing the predicitions, we again computed a posterior distribution, this time using the previously found posterior distribution (from the mean 0.8 and sd 0.2 prior and the original data) for each of the teachers as the new updated prior. Here, we used the new data for calculating the likelihood.

Then, we found the highest posterior density intervals (HPDI) at 50 percent by sampling from the posterior. Results are shown in the following table and plots: 
